Modelling the gas kinematics of an atypical Lyman-$\alpha$ emitting compact dwarf galaxy

### 2017-03-13 paper/letter.tex
new referencies

This theoretical prediction was only confirmed thirty years later on
distant, relatively young, not primordial, galaxies \citep{1998ApJ...498L..93D}.
Currently Lyman Alpha Emitting (LAE) galaxies are commonly targeted
in surveys. 
The presence of the Ly-$\alpha$ emission line provides confirmation of
the distance of a galaxy while provides clues about the stellar
population and inter-stellar medium conditions regulating the
Ly-$\alpha$ emission.

The Ly-$\alpha$ emission line is not exclusive of distant galaxies. 
Any galaxy with low dust content and ongoing star formation has the
right conditions to show this line.  
There are, for instance,  local Universe surveys that target
Ly-$\alpha$ emission in nearby dwarf star forming galaxies 
\citep{LARS}. 
The study of nearby LAE samples has allowed the study of other
indicators that might be more difficult to obtain for distant galaxies
such as morphology, dust attenuation, neutral hydrogen contents and
ionization state.

However, the physical interpretation of Ly-$\alpha$ observations is
not straightforward \citep{2015ApJ...805...14R}. 
This is due to the resonant nature of the Ly-$\alpha$ line. 
A Ly-$\alpha$ photon follows a diffusion-like process before escaping
the galaxy or being absorbed by dust. 
The resulting line profile becomes sensitive to the dynamical, chemical
and thermal conditions in the interstellar medium. 
There are very few analytically tools available to interpret the
emerging Ly-$\alpha$ line.
They are applicable only in very few cases of highly symmetrical
conditions, which are hardly met in real astrophysical systems.
For these reasons the interpretation of Ly-$\alpha$ observations
requires state-of-the-art Monte Carlo radiative transfer simulations.

\tol\ is a compact star forming dwarf galaxy that presents a
strong Ly-$\alpha$ emission \citep{Thuan97} with two puzzling 
features: the line is symmetric and single peaked.
Usually the Ly-$\alpha$ line has an asymmetric single or double peak. 
These two special features in \tol\ cannot be explained with
conventional models \citep{2006A&A...460..397V,2015ApJ...812..123G}.

>>>>

This theoretical prediction was only confirmed thirty years later on
distant, relatively young, not primordial, galaxies \citep{1998ApJ...498L..93D}.
Currently Lyman Alpha Emitting (LAE) galaxies are commonly targeted
in surveys. 
The presence of the Ly-$\alpha$ emission line provides confirmation of
the distance of a galaxy while provides clues about the stellar
population and inter-stellar medium conditions regulating the
Ly-$\alpha$ emission.
A careful clustering analysis of LAEs can also yield clues about their link
to dark matter halos \cite{2016ApJ...828....5M}.

The Ly-$\alpha$ emission line is not exclusive of distant galaxies. 
There are local Universe surveys that target Ly-$\alpha$ emission in nearby dwarf star forming galaxies  \citep{LARS}. 
The study of nearby LAE samples has allowed the study of other
indicators that might be more difficult to obtain for distant galaxies
such as morphology, dust attenuation, neutral hydrogen contents and
ionization state.

However, the physical interpretation of Ly-$\alpha$ observations is
not straightforward \citep{2015ApJ...805...14R}. 
This is due to the resonant nature of the Ly-$\alpha$ line. 
A Ly-$\alpha$ photon follows a diffusion-like process before escaping
the galaxy or being absorbed by dust. 
The resulting line profile becomes sensitive to the dynamical, chemical
and thermal conditions in the interstellar medium. 
There are very few analytically tools available to interpret the
emerging Ly-$\alpha$ line
\cite{Harrington73,1991ApJ...370L..85N,LoebRybicki,2006ApJ...645..792T}. 
They are applicable only in very few cases of highly symmetrical
conditions, which are hardly met in real astrophysical systems.
For these reasons the interpretation of Ly-$\alpha$ observations
requires state-of-the-art Monte Carlo radiative transfer simulations.

\tol\ is a compact star forming dwarf galaxy that presents a
strong Ly-$\alpha$ emission \citep{Thuan97} with two puzzling 
features: the line is symmetric and single peaked.
Usually the Ly-$\alpha$ line has an asymmetric single or double peak. 
These two special features in \tol\ cannot be explained with
conventional shell models
\citep{2006A&A...460..397V,2015A&A...578A...7V, 2015ApJ...812..123G}.
